---
title: "STA302 Final Project"
output: pdf_document
fontsize: 12pt
---



```{r,message=FALSE, echo=FALSE,warning=FALSE, include=FALSE}
library(tidyverse) 
library(latex2exp)
library(gridExtra)
data=read_csv("insurance.csv")
dataclean= na.omit(data)
selected <- dataclean[sample(1:nrow(dataclean), 1000, replace=F), ]
set.seed(1)
split <- sample(1:1000, 500, replace=FALSE) 
train<- selected[split,]
test <- selected[-split,]
```



The important variables are summarized in the following table.\
```{r,echo=FALSE,meassage=FALSE,warning=FALSE}
sum=dataclean[,c(1,3,6, 7)]
knitr::kable(summary(sum),caption = "Summary Table of the important variable")
```



```{r,message=FALSE, echo=FALSE,warning=FALSE}
attach(train)
par(mfrow=c(2,2))
hist(data$charges,col="pink",main="individual medical cost")
hist(log(data$charges),col="pink",main="distribution of medical log cost")
ggplot(data) +
   ggtitle("Plot of frequencies of charges categoried by smoker")+
    geom_histogram(aes(x =charges,fill = factor(smoker)), colour = "black", bins = 50,position = "dodge") +
    labs(y = "Frequency", x = "charges")

library(ggplot2)
ggplot(data, aes(age, charges)) +
  geom_smooth(se = FALSE,method="lm", formula = 'y ~ x') +
  geom_point(aes(color = smoker))+
  labs(
    y = "medical expenditure", x = "age",
    title = "relation between medical expense and age",
    subtitle = "categoried by smoker"


  )

ggplot(data, aes(children, charges)) +
 geom_smooth(se = FALSE,method="lm", formula = 'y ~ x') +
  geom_point(aes(color = sex))+
  labs(
    y = "medical expenditure", x = "children number",
    title = "relation between medical expense and children number",
    subtitle = "categoried by gender"

  )

ggplot(data,aes(bmi, charges)) +
  geom_smooth(se = FALSE,method="lm", formula = 'y ~ x') +
  geom_point(aes(color = region))+
  labs(
    y = "medical expenditure", x = "Body mass index",
    title = "relation between medical expense and gender",
    subtitle = "categoried by region"

  )


```



```{r,message=FALSE, echo=FALSE,warning=FALSE}
library(kableExtra)
m1=lm(charges~age+sex+bmi+smoker+region+children ,data=train)
kbl(broom::tidy(m1)[,], caption = "Summary Table of Model 1", 
      digits = 3, format = "markdown", align = "ccccc", padding= 3) 

m10=lm(charges~age ,data=train)
knitr::kable(anova(m1,m10),cap="ANOVA TABLE For partial F test")
r <- resid(m1)
# first check condition 1 and 2
#condition 1
plot(charges ~ fitted(m1), main="Y versus Y-hat", xlab="Y-hat", ylab="Y")
abline(a = 0, b = 1)
lines(lowess(charges ~ fitted(m1)), lty=2)
#condition 2
data2 = data.frame(train$charges, train$age, train$bmi)
pairs( data2 )
```
```{r,message=FALSE, echo=FALSE,warning=FALSE}
train$charges <- log(train$charges)
m2=lm(charges~age+sex+bmi+smoker+region+children ,data=train)
kbl(broom::tidy(m2)[,], caption = "Summary Table of Model 2", 
      digits = 3, format = "markdown", align = "ccccc", padding= 3) 

```

```{r,message=FALSE, echo=FALSE,warning=FALSE}
train$charges <- log(train$charges)
m3=lm(charges~age+smoker+region+children ,data=train)
kbl(broom::tidy(m3)[,], caption = "Summary Table of Model 3", 
      digits = 3, format = "markdown", align = "ccccc", padding= 3) 
m10=lm(charges~age ,data=train)
knitr::kable(anova(m3,m10),cap="ANOVA TABLE For partial F test")
r <- resid(m3)
# first check condition 1 and 2
#condition 1
plot(charges ~ fitted(m3), main="Y versus Y-hat", xlab="Y-hat", ylab="Y")
abline(a = 0, b = 1)
lines(lowess(charges ~ fitted(m3)), lty=2)
#condition 2
data2 = data.frame(train$charges, train$age, train$bmi)
pairs( data2 )

library(car)

plot(rstandard(m3)~fitted(m3), xlab="fitted", ylab="Residuals")

qqnorm(rstandard(m3))
qqline(rstandard(m3))
```


```{r,message=FALSE, echo=FALSE,warning=FALSE}
m4=lm(charges~ bmi + children + region ,data=train)
kbl(broom::tidy(m4)[,], caption = "Summary Table of Model 4", 
      digits = 3, format = "markdown", align = "ccccc", padding= 3) 
```




```{r,message=FALSE, echo=FALSE,warning=FALSE}
m10=lm(charges~age ,data=train)
knitr::kable(anova(m4,m10),cap="ANOVA TABLE For partial F test")
r <- resid(m4)
# first check condition 1 and 2
#condition 1
plot(charges ~ fitted(m4), main="Y versus Y-hat", xlab="Y-hat", ylab="Y")
abline(a = 0, b = 1)
lines(lowess(charges ~ fitted(m4)), lty=2)
#condition 2
data2 = data.frame(train$charges, train$children, train$bmi)
pairs( data2 )

library(car)

plot(rstandard(m4)~fitted(m4), xlab="fitted", ylab="Residuals")

qqnorm(rstandard(m4))
qqline(rstandard(m4))
```




```{r,message=FALSE, echo=FALSE,warning=FALSE, include=FALSE}
#find outlier
r <- rstandard(m4)
out <- which(r > 5 | r < -5)
out
# determine whether there are leverage points
h <- hatvalues(m4)
threshold <- 2 * ((length(m4$coefficients)+1)/nrow(train))
w <- which(h > threshold)
train[w,]
# find all influence functions
D <- cooks.distance(m1)
cutoff <- qf(0.5, length(m1$coefficients), nrow(train)-length(m1$coefficients), lower.tail=T)
which(D > cutoff)
fits <- dffits(m4)
library(car)
vif(m4)
```


```{r,message=FALSE, echo=FALSE,warning=FALSE}
#model checking
test$charges <- log(test$charges)
m5=lm(charges~bmi+region+children ,data=test)
kbl(broom::tidy(m2)[,], caption = "Summary Table of Model 3", 
      digits = 3, format = "markdown", align = "ccccc", padding= 3) 

```
```{r,message=FALSE, echo=FALSE,warning=FALSE}
r <- resid(m5)
# first check condition 1 and 2
#condition 1
plot(charges ~ fitted(m5), main="Y versus Y-hat", xlab="Y-hat", ylab="Y")
abline(a = 0, b = 1)
lines(lowess(charges ~ fitted(m5)), lty=2)
#condition 2
data2 = data.frame(test$charges, test$children, test$bmi)
pairs( data2 )
plot(rstandard(m5)~fitted(m5), xlab="fitted", ylab="Residuals")
qqnorm(rstandard(m5))
qqline(rstandard(m5))
```





```{r}

select = function(model, n)
{
  SSres <- sum(model$residuals^2)
  Rsq <- summary(model)$r.squared
  Rsq_adj <- summary(model)$adj.r.squared
  p <- length(model$coefficients) - 1
  AIC <- n*log(SSres/n) + 2*p     # you could also use AIC()
  AICc <- AIC + (2*(p+2)*(p+3)/(n-p-1))
  BIC <- n*log(SSres/n) + (p+2)*log(n)    # could also use BIC()
  res <- c(SSres, Rsq, Rsq_adj, AIC, AICc, BIC)
  names(res) <- c("SSres", "Rsq", "Rsq_adj", "AIC", "AIC_c", "BIC")
  return(res)
}

# apply to the models
round(select(m1, nrow(train)))
round(select(m2, nrow(train)))
round(select(m3, nrow(train)))
round(select(m4, nrow(train)))
round(select(m5, nrow(test)))



```


```{r}

```

Running this function on each model would allow us to extract the values we wish to display in a table that would summarize the results of our model selection approach.

Model | Adjusted $R^2$ | AIC | BIC 
------|----------------|-----|-----
Full model | `r round(s1[3], 2)` | `r s1[4]` | `r s1[6]`
- Birdie Conversion and Putts Per round | `r s3[3]` | `r s3[4]` | `r s3[6]`
- Putting Average and Putts Per round | `r s5[3]` | `r s5[4]` | `r s5[6]`





